---
title: "Exercise sheet 2"
date: "Nov 2025"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Exercise 1
1. Create a vector `vec` containing _each_ of
numbers of the sequence from 22 to 35 (in steps of 3) 4 times!

2. Print the length of `vec` in the format "vec has length LENGTH" using the `length` and `paste` functions. _Tip:_ Use the `?` function and Google first if you can't remember how to use `paste`.

3. Create another vector `vec2` of length 5 using any of the functions from the lecture.

4. Add `vec` and `vec2`. What happens?
_Tip:_ Look at which numbers have been added together!

5. What happens if the vectors are of different lengths?

## Exercise 2
1. Print out the 1st, 5th and 8th elements of `vec`.

2. Print out all elements of `vec` _except_ for elements 1, 5 and 8.

3. Using the `which` function, find the indices of all elements...
  - ...larger than 25.
  - ...smaller than 28 but larger than 22. _Tip:_ Combine conditions using the `&` (logical AND) operator!
  - ...which are divisible by 3. _Tip:_ A value `x` is divisible by `a` if `x %% a == 0`.

## Exercise 3
1. Create a new vector `log_vec` containing the values of `vec` after applying the `log` function.

2. Round `log_vec`...
  - ...down.
  - ...up.
  - ...to 2 digits.
  
3. What are the largest and smallest values of `log_vec`?

4. Use the equivalence operator `==` to compare `exp(log_vec)` to `vec`. What is printed out?

## Exercise 4
1. Create a new vector `vec3` containing all whole numbers from 1 to 500.

2. Use the `sample` function to draw 50 random numbers from `vec3` and save the result in a new vector `dat`. You can pick yourself if you want to draw with or without replacement!
_Tip:_ Use the `?` function and Google to find out how `sample` works!
An example use: 
```{r eval = FALSE}
candies <- c("licorice", "fudge", "gummy bears", "chocolate")
dat <- sample(candies, 20, replace = FALSE)
```

3. Calculate the mean of `dat`.

4. Calculate the standard deviation of `dat`.