---
title: "Exercise sheet 3"
author: "Johan Hennert"
date: "2023-11-15"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

# Exercise 1
## 1.1
1. Pick 4 cities and find\ldots 
  - the country they're in,
  - their area (in km\textsuperscript{2}), and
  - their population.



  
2. Create a data frame `cities` with four columns: `name`, `country`, `area`, `population`.

## 1.2
Find\ldots

- \ldots the area of the second city.

- \ldots the value in the 1st row and 3rd column.

- \ldots the entire 4th row.

## 1.3
Subset `cities` to\ldots

- \ldots get all cities with population above 1 million.

- \ldots find only the names of the cities with an area smaller than 200 km\textsuperscript{2}.

## 1.5
Add a new column `density` containing the population density of each city ($D_{ensity}=\frac{P_{opulation}}{A_{rea}}$).


## 1.6 
Add another new column `is_city_state` that is `TRUE` if the city and country names are equal,
and `FALSE` otherwise.

_Tip:_ The `==` operator also returns a vector!

_Bonus task:_ Try using the `ifelse` function to have "yes" and "no" instead of `TRUE` and `FALSE`.

\newpage
  
# Exercise 2
## 2.1
The `readline` function allows the user to enter some input and is used in the following way:
```{r 2.1_demo, echo = TRUE, eval = FALSE}
input <- readline("Enter something! ")
print(input) # print whatever was input
```

Write a code chunk where the user is asked for a name, and output a string containing "Hello, NAME!", with NAME being whatever the user gave as a name.

_Note:_ When using RMarkdown and executing a code chunk containing `readline` the input is given in the console (so check there if it looks
like nothing is happening).


## 2.2
Extend the code from the previous exercise to check if the given name is the same as your name. If this is the case, print "Hey, that's my name as well!", otherwise, print the same thing as in 2.1.


## 2.3
The `runif(n, min, max)` can be used to generate `n` random numbers between `min` and `max`^[Strictly defined `the runif` function generates `n` numbers from a _uniform distribution_ between `min` and `max`. This means, each number between `min` and `max` is equally likely to be chosen.].

1. Use it to generate one random number between 1 and 10.
2. Have the user enter a number using `readline`.
3. Check if the guessed number was the randomly generated number.
  - If it was, print "Correct!",
  - otherwise print "The correct number was NUMBER", where NUMBER is the generated number.

_Note:_ `runif` will generate numbers with a lot of digits, so to make it easier for the guesser, round the random number to 0 digits.


## 2.4
Create a vector from 40 to 60 and use it together with the `ifelse` function to get
a vector with values "Too small!" for anything less than 50, "Just right!" for everything greater than
or equal to 50 and less than 60, and "Too large!" for numbers greater than or equal to 60.